---
title: "ex1"
output:
  pdf_document: default
  html_document: default
---
##id:30788595-4

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
#function
```{r}

bisection<- function(fun=fun, digit = 10,a,b, step = F)
{
  if ((fun(a)>=0 & fun(b)>=0) || (fun(a)<0 & fun(b)<0) ) {
    print("A and B have an equal sign")
    return()
  }
  while ((b-a)>(10^-digit)) {
    c <- (a+b)/2
     if (step) {
      print(c)
    }
    if (fun(c)==0) {
      return(c)
    }else if (fun(a)*fun(c)>0) {
      a <- c
    }else {
      b <- c
    }
  }
  return(c)
 
}

Newton <- function(fun=fun, digit = 10,x, step = F){
  x_1 <- fun(x)
  while (round(x,digits = digit)!=round(x_1,digits = digit)) {
    if (step) {
      print(x_1)
    }
    x <- x_1
    x_1<- fun(x_1)
  }
return(x_1)
}


```

#Q1
```{r}
library(tictoc)
tic()
f_1 <- function(x){2*sin(x)-x}
bisection(fun = f_1,digit = 6,0.5*pi,pi,step = F)
toc()

tic()
f_2 <- function(x){x - (2*sin(x)-x)/(2*cos(x)-1)}
Newton(fun = f_2,digit = 6,x = 0.5*pi,step = F)
toc()
```
The Newton method makes fewer iterations, as expected, but is not always faster. In each run, different results are obtained, and sometimes it is turns out that the bisection method is faster. The result is likely to come from R, and if we write in C or a more basic programming language we will get results that match that logic.


#Q2
```{r}

tic()
f <- function(x) {x*(2-sqrt(2)*x)}
Newton(fun = f,digit = 6,x = 1)
toc()

```
$$f(x)= 1/x - \sqrt2$$
$$f'(x) = -/x^2$$
$$x_1 = x_0 - \frac{-1/x_0 - \sqrt2}{-1/x^2}$$
$$x_0 + x_0^2 (1/x_0 - \sqrt{2})$$
$$2x-\sqrt{2}x^2$$
$$x_1=x_0(2-\sqrt{2}x_0)$$

Because the function has one root, if the algorithm converged to a solution, it converged to the only solution.

#Q3
```{r}
tic()
f_1 <- function(x){x - 1*((exp(x)-x-1)/(exp(x)-1))}
Newton(fun = f_1,digit = 6,x = 1)
toc()

tic()
f_2 <- function(x){x - 2*((exp(x)-x-1)/(exp(x)-1))}
Newton(fun = f_2,digit = 6,x = 1)
toc()

```

The derivative is equal to: $e^x-1=0$
It is easy to see that there is a single root at X = 0


#Q4
```{r}

f <- function(x){x - atan(x)*((x^2)+1)}
Newton(fun = f,digit = 6,x = -1)

```


$X_{n+1} = X_n - (1+X^2 _n)arctan(x_n)$
The Newton formula is used to find a local minimum. If $X_0$ is selected too large, you can see that $X_{n+1}$ will be larger and the formula will start to explode.
The way to find the solution is to place $X_n$ and $X_{n+1}$ the $X*$. We will receive:
$f(X*) = 2x* - (1+x*^2) arctan(x*)$
It is not hard to estimate $x* = 1.3917 $.
